---
title: "Week2"
date: "2/13/2022"
output: html_document
---

## Step 1: Downloading, uncompressing and loading data

The following code lines download the file activity.zip from the url provided and unzip the file. 

```{r echo=TRUE}
# Defining the urlprovided from the task description:
urlProvided <-   "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
# Downloading the file 
download.file(urlProvided, destfile = "activity.zip")
# Unzip 
unzip("activity.zip")
# load data
dataBase <- read.csv("activity.csv")
# Convert date to date format
dataBase$date <- as.Date(dataBase$date)
head(dataBase)
```
## Step 2: Histogram of the total number of steps taken each day

Sum steps per day, then plot using hist. 

```{r echo=TRUE}
stepsperday <- aggregate(dataBase$steps, list(Day = dataBase$date), FUN = sum )
hist(stepsperday$x, breaks = 20, main = "Steps taken each day", xlab = "Steps by day")
```

## Step 3: Mean and median number of steps taken each day

```{r echo=TRUE}
mean(stepsperday$x, na.rm = TRUE)
median(stepsperday$x, na.rm = TRUE)
```

## Step 4: Time series plot of the average number of steps taken
```{r echo=TRUE}
filt_dataBase <- dataBase[ is.na(dataBase$steps) != TRUE , ]
stepsperinterval <- with(filt_dataBase, aggregate(steps, list(interval), FUN = mean ))
with(stepsperinterval, plot(Group.1, x, type = "l", xlab = "Interval", ylab = "Steps"))

```
## The 5-minute interval that, on average, contains the maximum number of steps

##Code to describe and show a strategy for imputing missing data

## Histogram of the total number of steps taken each day after missing values are imputed

##Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

All of the R code needed to reproduce the results (numbers, plots, etc.) in the report